---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# melt - Multiple Empirical Likelihood Tests
<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)
[![R-CMD-check](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Overview
The R package **melt** provides a unified framework for data analysis with empirical likelihood methods. A collection of functions are available for basic regression analysis and hypothesis testing. Much of its functionality and syntax are designed to mimic the corresponding base R functions. The core routines are written in C++ and utilize OpenMP for parallelization. 


## Installation
You can install the latest stable release from [CRAN](https://cran.r-project.org/package=melt).
```{r eval = FALSE}
install.packages("melt", dependencies = TRUE)
```
You can install the latest development version from
[Github](https://github.com/markean/melt).
```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("markean/melt")
```


## Usage
```{r eval = TRUE}
library(melt)
# one sample test (mean)
el_mean(par = 0, x = rnorm(n = 100))

# linear regression
fit <- el_lm(formula = mpg ~ wt, data = mtcars)
summary(fit)

# analysis of variance
el_aov(formula = Sepal.Length ~ Species, data = iris)
```
