---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# melt - Multiple Empirical Likelihood Tests
<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/melt)](https://CRAN.R-project.org/package=melt)
[![R-CMD-check](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/markean/melt/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/markean/melt/branch/master/graph/badge.svg)](https://app.codecov.io/gh/markean/melt?branch=master)
<!-- badges: end -->

## Overview
The R package **melt** provides a unified framework for data analysis with empirical likelihood methods. A collection of functions are available for regression analysis and hypothesis testing. Much of its functionality and syntax mimics the corresponding base R functions. The core computational routines are implemented with the 'Eigen' C++ library and 'RcppEigen' interface, with OpenMP for parallel computation. Additional functions are available for multiple testing for the analysis of experimental designs. Details of the testing procedures are given in [Kim et al. (2021)](https://arxiv.org/abs/2112.09206). 


## Installation
You can install the latest stable release from [CRAN](https://cran.r-project.org/package=melt).
```{r eval = FALSE}
install.packages("melt")
```
You can install the latest development version from
[Github](https://github.com/markean/melt).
```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("markean/melt")
```


## Usage
```{r eval = TRUE}
library(melt)
# one sample test for mean
fit1 <- el_mean(par = 0, x = rnorm(n = 100))
confint(fit1)


# linear regression
fit2 <- el_lm(formula = mpg ~ wt, data = mtcars)
summary(fit2)
confint(fit2)


# analysis of variance 
data("clothianidin")
fit3 <- el_lm(clo ~ -1 + trt, clothianidin)
summary(fit3)
confint(fit3)


# test of no treatment effect
lhs <- matrix(c(1, -1, 0, 0,
                0, 1, -1, 0,
                0, 0, 1, -1), byrow = TRUE, nrow = 3)
elt(fit3, lhs = lhs)


# logistic regression
fit4 <- el_glm(vs ~ mpg, data = mtcars, family = binomial)
summary(fit4)
```
